### 2017-05-08 2017mergingmerges/section/02-Processes.tex
Appendix: finite stream rules

The @group@ combinator has two parameters, @sIn1@ and @sOut1@, which bind the input and output streams respectively. The \emph{nu-binders} \mbox{$\nu$ @(f: Bool) (l: Nat)@...} indicate that each time the @group@ combinator is instantiated, fresh names must be given to @f@, @l@ and so on, that do not conflict with other instantiations. Overall, the @f@ variable tracks whether we are dealing with the first value from the stream, @l@ holds the last value pulled from the stream (or 0 if none have been read yet), and @v@ holds the current value pulled from the stream.

@@

Our fusion algorithm takes two processes and produces a new one that computes the output of both. For example, suppose we need a single process that produces the output of the first two lines of our @uniquesUnion@ example back in \S\ref{s:Introduction}. The result will be a process that computes the result of both @group@ and @merge@ as if they were executed concurrently, where the first input stream of the @merge@ process is the same as the input stream of the @group@ process. When describing the fusion algorithm we will instantiate the parameters of each combinator with arguments of the same names.

@@

The algorithm proceeds by considering pairs of states: one in each of the source process state machines to be fused. Both the @group@ machine and the @merge@ machine pull from the same stream as their initial instruction, so we have the situation shown in Figure~\ref{fig:Fusion:Pulls}. The @group@ machine needs to transition from label @A0@ to label @A1@, and the @merge@ machine from @B0@ to @B1@. In the result machine we produce three new instructions that transition between four joint result states, @F0@ to @F3@.
Each of the joint result states represents a combination of two source states, one from each of the source machines. For example, the first result state @F0@ represents a combination of the @group@ machine being in its initial state @A0@ and the @merge@ machine being in its own initial state @B0@.

We also associate each of the joint result states with an description of whether each source machine has already pulled a value from each of its input streams. For the @F0@ case shown in Figure~\ref{fig:Fusion:Pulls} we have ((A0, \{sIn1 = none\}), (B0, \{sIn1 = none, sIn2 = none\})). The result state @F0@ represents a combination of the two source states @A0@ and @B0@. As both @A0@ and @B0@ are the initial states of their respective machines, those machines have not yet pulled any values from their two input streams, so both `sIn1' and `sIn2' map to `none'.

@@

% -----------------------------------------------------------------------------
\subsubsection{Fusing Cases}
Once the result machine has arrived at the joint state @F3@, this is equivalent to the two source machines arriving in states @A1@ and @B1@ respectively. The left of Figure~\ref{fig:Fusion:Case} shows the next few transitions of the source machines. From state @A1@, the @group@ machine needs to perform a @case@ branch to determine whether to push the current value it has from its input stream @sIn1@ to output stream @sOut1@, or to just pull the next value from its input. From state @B1@, the @merge@ machine needs to pull a value from its second input stream @sIn2@. In the result machine, @F3@ performs the case analysis from @A1@, moving to either @A2@ or @A3@, corresponding to @F4@ and @F5@ respectively. From state @F4@, the push at @A2@ is executed and moves to @A3@, corresponding to @F5@. Finally, at @F5@ the @merge@ machine pulls from @sIn2@, moving from @B1@ to @C0@. Because the stream @sIn2@ is only pulled from by the @merge@ machine, no coordination is required between the @merge@ and @group@ machines for this pull.

>>>>

The @group@ combinator has two parameters, @sIn1@ and @sOut1@, which bind the input and output streams respectively. The \emph{nu-binders} \mbox{($\nu$ @(f: Bool) (l: Nat)@...)} indicate that each time the @group@ combinator is instantiated, fresh names must be given to @f@, @l@ and so on, that do not conflict with other instantiations. Overall, the @f@ variable tracks whether we are dealing with the first value from the stream, @l@ holds the last value pulled from the stream (or 0 if none have been read yet), and @v@ holds the current value pulled from the stream.

@@

Our fusion algorithm takes two processes and produces a new one that computes the output of both. For example, suppose we need a single process that produces the output of the first two lines of our @uniquesUnion@ example back in \S\ref{s:Introduction}. The result will be a process that computes the result of both @group@ and @merge@ as if they were executed concurrently, where the first input stream of the @merge@ process is the same as the input stream of the @group@ process. In our informal description of the fusion algorithm we instantiate the parameters of each combinator with arguments of the same names.

@@

The algorithm proceeds by considering pairs of states: one from each of the source process state machines to be fused. Both the @group@ machine and the @merge@ machine pull from the same stream as their initial instruction, so we have the situation shown in the top of Figure~\ref{fig:Fusion:Pulls}. The @group@ machine needs to transition from label @A0@ to label @A1@, and the @merge@ machine from @B0@ to @B1@. In the result machine we produce three new instructions that transition between four joint result states, @F0@ to @F3@.
Each of the joint result states represents a combination of two source states, one from each of the source machines. For example, the first result state @F0@ represents a combination of the @group@ machine being in its initial state @A0@ and the @merge@ machine being in its own initial state @B0@.

We also associate each of the joint result states with a description of whether each source machine has already pulled a value from each of its input streams. For the @F0@ case in the top of Figure~\ref{fig:Fusion:Pulls} we have ((A0, \{sIn1 = none\}), (B0, \{sIn1 = none, sIn2 = none\})). The result state @F0@ represents a combination of the two source states @A0@ and @B0@. As both @A0@ and @B0@ are the initial states of their respective machines, those machines have not yet pulled any values from their two input streams, so both `sIn1' and `sIn2' map to `none'.

@@

% -----------------------------------------------------------------------------
\subsubsection{Fusing Cases}
Once the result machine has arrived at the joint state @F3@, this is equivalent to the two source machines arriving in states @A1@ and @B1@ respectively. The lower half of Figure~\ref{fig:Fusion:Case} shows the next few transitions of the source machines. From state @A1@, the @group@ machine needs to perform a @case@ branch to determine whether to push the current value it has from its input stream @sIn1@ to output stream @sOut1@, or to just pull the next value from its input. From state @B1@, the @merge@ machine needs to pull a value from its second input stream @sIn2@. In the result machine, @F3@ performs the case analysis from @A1@, moving to either @A2@ or @A3@, corresponding to @F4@ and @F5@ respectively. From state @F4@, the push at @A2@ is executed and moves to @A3@, corresponding to @F5@. Finally, at @F5@ the @merge@ machine pulls from @sIn2@, moving from @B1@ to @C0@. Because the stream @sIn2@ is only pulled from by the @merge@ machine, no coordination is required between the @merge@ and @group@ machines for this pull.
